Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-29 paper.tex
add discussion about time step duration effects on performance

Another difference between fleet-based and individual facility modeling
involves fuel sharing.  Power production by a group of individual reactors for
a fixed amount of fuel shortage may vary depending on how the existing fuel is
distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
being short 3 batches might mean that one reactor has no batches and isn't
operating or that 3 reactors are each missing a single batch.  This is
referred to as the \emph{fuel sharing effect}.  Fleet reactors, by design,
have perfect fuel sharing - all fuel is distributed so that the minimum
possible amount of capacity is offline during shortages. The most correct
measure of fuel sharing inefficiency is the difference between the actual
generated power and the maximum amount of power that could have been generated
among all possible fuel distribution alternatives.  This is difficult to
actually compute. One possible approximation is to count the number of fresh
fuel batches distributed to reactors that ended up not having enough fuel to
operate (i.e. a full core) that time step.  This approximation assumes that
every "wasted" batch of fuel, could have been given to a reactor that only
needed one batch to operate.

@@

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps and individual reactor modeling
create a need for larger floating inventory buffers in order to avoid material
shortages.  Another way to think about this effect is that a single material
object can only move between a single pair of facilities per time step. Larger
time steps result in longer minimum bounds on time to traverse paths between
facilities. Although direct measures of this effect are difficult, some of its
consequences can be quantified by comparing scenarios with different time step
durations (e.g. cases 2 and 4) such as higher standing inventories during
shortages.

There are also performance differences between the two modeling paradigms.  A
single 200 year simulation with several hundred fleet-based reactors takes
about 2 seconds to run on a typical desktop machine where the same simulation
using individual reactor facilities takes about 20 seconds.  This order of
magnitude difference is an important tradeoff to keep in mind when doing fuel
cycle analysis in general, particularly when considering optimization or
sensitivity studies where running hundreds of thousands of simulations may be
desirable.

>>>>

Another difference between fleet-based and individual facility modeling
involves fuel sharing.  Power production by a group of individual reactors for
a fixed amount of fuel shortage may vary depending on how available fuel is
distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
being short 3 batches might mean that one reactor has no batches and isn't
operating or that 3 reactors are each missing a single batch.  This is
referred to as the \emph{fuel sharing effect}.  Fleet reactors, by design,
have perfect fuel sharing - all fuel is distributed so that the minimum
possible amount of capacity is offline during shortages. The most correct
measure of fuel sharing inefficiency is the difference between the actual
generated power and the maximum amount of power that could have been generated
among all possible fuel distribution alternatives.  This is difficult to
actually compute. One possible approximation is to count the number of fresh
fuel batches distributed to reactors that ended up not having enough fuel to
operate (i.e. a full core) that time step.  This approximation assumes that
every "wasted" batch of fuel, could have been given to a reactor that only
needed one batch to operate.

@@

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps create a need for larger floating
inventory buffers in order to avoid material shortages.  Another aspect of
this effect is that a single material object can only move between a single
pair of facilities per time step. Larger time steps result in longer minimum
bounds on time to traverse paths between facilities. Although direct measures
of this effect are difficult, some of its consequences (e.g. higher standing
inventories during shortages) can be quantified by comparing scenarios with
different time step durations (e.g. cases 2 and 4).

There are also performance differences between the two modeling paradigms.  A
single 200 year simulation with several hundred fleet-based reactors takes
about 2 seconds to run on a typical desktop machine where the same simulation
using individual reactor facilities takes about 20 seconds.  This order of
magnitude difference is an important tradeoff to keep in mind when doing fuel
cycle analysis in general, particularly when considering optimization or
sensitivity studies where running hundreds of thousands of simulations may be
desirable.  The time step duration has less impact on performance because the
amount of work being done in the simulation is the same; in the monthly time
step cases most facilities (i.e. all reactors) are only perform actions every
refueling cycle rather than every time step.
